Demo for a Quality of Service networking with Software Defined Network.

### 2017-03-15 report/paper.tex
final correction. matteo need to add amount of gaming traffic he analyzed

\textbf{Forwarding table}
\begin{itemize}
 \item \texttt{match fields:} compare packets properties in order to look for correspondences. They contain the in_port, the packet's header and others optional fields;
 \item \texttt{priority:} this field is used when the specific packet matches more than one rule. The higher the number, the higher the priority a rule has;
 \item \texttt{counter:} this field was improved every time a correspondence with a packet is matched;
 \item \texttt{istruction:} a set of actions that the switch add on \emph{Action-set} of the matched packet or some instruction 
 in order to modify the pipeline flux;
 \item \texttt{timeout:} this is the maximum time afther which the rule expire;
 \item \texttt{flag:} with this field we can differenciate similar voices in order to managed multiple table with similar rules;
 \\
 \\
\end{itemize}

@@

\subsection{Known Limitations}
SDN networks are known to be underperforming with low traffic, which means that if nodes tries to connect with each other
to exchange little traffic, the network will have a higher delay and will perform worst than a
traditional network. We tried to reduce this underperformance with the usage of the wildcards:
Every rule needs at least the source ip or the destination ip. If one of them is in a network
(e.g. the internet traffic is in the network 0.0.0.0/0 excluded 192.168.10.0/24 ),
we can redirect all the traffic from an internal host to the internet with a predefined
path. This means that only one PacketIn will be sent to the controller if a host is surfing the network.

Another limitation is the no default routing from a source to a destination: when an host
is plugged to a network of SDN switches, it won't know the destination. The only way it can
discover the host is by flooding the network with that packet, but this makes each switch raise a PacketIn message to the controller
slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
the network topology, but doesn't know how much traffic each host is doing. For that the administrator need to put a firewall, which it's not needed in SDN networks,
as the controller can set a switch to drop all the packets from/to a destination.

@@

\subsection{Adaptability of the network}
An adaptable Network has to real-time reroute traffic in order to ensure a lot of demands.
A general efficient Network has to manage some tasks, for example avoiding links congestion, reroute traffic when some malfunction occure on a link
and running a link path algorithm. 
In addition our network has to be adaptable in order to ensure a QoS type of service.

@@

\subsection{Exterior comunication}
Our network needed to be able to communicate with the external internet to be useful for a demo which demonstrate the capabilities of a SDN network.
The first thing we needed to accomplish was to make the controller aware of what switch and which port of the switch is connected to the external internet:
to do so, we allowed our controller to be able to receive and process in-line parameters. We provided two parameters to the controller, so that
it could recognize where the traffic to the internet need to be sent. These two parameters were the mac address of the
port of a switch that is connected to the link and the network address of our network, to let the controller know if a request from a host was meant to another host
within the network or the data stream was intended to the internet. We provided only one gateway, as the Routeboard had few ports available for the packet switching
and only four SDN-capable switches to be able to simulate a real network that was generating traffic.

@@

We have use 4 MikroTik Routerboards that were provided us by the Communication's Network Lab in the department of information  engineering to 
University of Padua.
The 4 routerboards are connected each other in order to design a topological loop. Then every sigle board connects two hosts and is connected with a 
legacy switch which provides logically connection with controller.
The last thing is a host that made gateway's function; we have connected its with both routerboard and legacy network and set up this host like default 
gateway on the controller.

\subsection{Graph and Network's structures}
  In order to achieve the realtime adaptability and statistics storage we developed ad-hoc structures.
  We used a specific Python's library called NetworkX \cite{networkx}. NetworkX allow us to create a graph in witch we can storage suitable nodes
  that represent resplectively switches or hosts.
  This topological structure is updated live thanks to connection events and links discover offered by POX's libraries:
  \begin{itemize}
   \item \texttt{_handle_LinkEvent}: from openflow.discovery an event of this type gives us two types of information; a new link is added and 
   a still present link is expired;
   \item \texttt{_handle_ConnectionUp}: from core.openflow this event joins up when a new switch is added on the network and gives us
   some information of the switch, in particular its ports configuration;
   \item \texttt{_handle_ConnectionUp}: from core.openflow like the event explained above comes up when a switch is disconnected from the network;
   \item \texttt{_handle_port_status}: from core.openflow this event occures when on a particular switch there is a change about its ports, for example
   a new host has just been connected or the link is expired;\\
  \end{itemize}

@@

\subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  hosts  or  middleboxes.
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  simplifies
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occur from sending a packet from controller to a switch and the time in which the packet return from
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifying the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculate RTT for the switch.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}
 \caption{\emph{RTT between controller and OF switch}}
 \label{fig:topo}
\end{figure}
\newline
If we need to calculate the delay on a specific link, the rtt has to calculate with the headed switch of this link.
\\
\newline
Now we create a proper message with unique id. number, let's start a timer \texttt{TM1} and then send this message to a first switch specifying the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particular flow entry on the second switch in order to send to the controller only some particular packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \caption{\emph{Pck from controller to itself following a particular way.}}
 \label{fig:topo}
\end{figure}
\\
\newline At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Theoretically this method must works but in practice there are a lot of delay introduced from computational congestion on the virtualized controller.
What we observe is a estimated delay like a few seconds which is not the case, in fact if we made some \texttt{ping test} from one host to another host
in this SDN network we realize delay around 3ms.
\\
\\
The following picture shows the visual output that we can obtain from delay's graph. In this graph we storage only the core information;
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.26]{images/delay.png}
 \caption{\emph{Delay's graph.}}
 \label{fig:topo}
\end{figure}

>>>>

\textbf{Forwarding table}
\begin{itemize}
 \item \texttt{match fields:} compare packets properties in order to look for correspondences. They contain the in_port, the packet's header and others optional fields;
 \item \texttt{priority:} this field is used when the specific packet matches more than one rule. The higher the number, the higher the priority a rule has;
 \item \texttt{counter:} this field was improved every time a correspondence with a packet is matched;
 \item \texttt{istruction:} a set of actions that the switch add on \emph{Action-set} of the matched packet or some instruction 
 in order to modify the pipeline flux;
 \item \texttt{timeout:} this is the maximum time after which the rule expire;
 \item \texttt{flag:} with this field we can differentiate similar voices in order to managed multiple table with similar rules;
 \\
 \\
\end{itemize}

@@

\subsection{Known Limitations}
SDN networks are known to be under performing with low traffic, which means that if nodes tries to connect with each other
to exchange little traffic, the network will have a higher delay and will perform worst than a
traditional network. We tried to reduce this under performance with the usage of the wildcards:
Every rule needs at least the source ip or the destination ip. If one of them is in a network
(e.g. the internet traffic is in the network 0.0.0.0/0 excluded 192.168.10.0/24 ),
we can redirect all the traffic from an internal host to the internet with a predefined
path. This means that only one PacketIn will be sent to the controller if a host is surfing the network.

Another limitation is the absence of the default routing from a source to a destination: when an host
is plugged to a network of SDN switches, it won't know the destination. The only way it can
discover the host is by flooding the network with that packet, but this makes each switch raise a PacketIn message to the controller
slowing down the startup of the connection. This thing doesn't happen with traditional routing, because each router knows
the network topology, but doesn't know how much traffic each host is doing. For that the administrator need to put a firewall, which it's not needed in SDN networks,
as the controller can set a switch to drop all the packets from/to a destination.

@@

\subsection{Adaptability of the network}
An adaptable Network has to real-time reroute traffic in order to ensure a lot of demands.
A general efficient Network has to manage some tasks, for example avoiding links congestion, reroute traffic when some malfunction occur on a link
and running a link path algorithm. 
In addition our network has to be adaptable in order to ensure a QoS type of service.

@@

\subsection{Exterior communication}
Our network needed to be able to communicate with the external internet to be useful for a demo which demonstrate the capabilities of a SDN network.
The first thing we needed to accomplish was to make the controller aware of what switch and which port of the switch is connected to the external internet:
to do so, we allowed our controller to be able to receive and process in-line parameters. We provided two parameters to the controller, so that
it could recognize where the traffic to the internet need to be sent. These two parameters were the mac address of the
port of a switch that is connected to the link and the network address of our network, to let the controller know if a request from a host was meant to another host
within the network or the data stream was intended to the internet. We provided only one gateway, as the Routeboard had few ports available for the packet switching
and only four SDN-capable switches to be able to simulate a real network that was generating traffic.

@@

We used 4 MikroTik Routerboards that were provided us by the Communication's Network Lab in the Department of Information Engineering to 
University of Padua.
The 4 routerboards are connected each other in order to design a topological loop. Then every single board connects two hosts and is connected with a 
legacy switch which provides logically connection with controller.
The last thing is a host that made gateway's function; we have connected its with both routerboard and legacy network and set up this host like default 
gateway on the controller.

\subsection{Graph and Network's structures}
  In order to achieve the real time adaptability and statistics storage we developed ad-hoc structures.
  We used a specific Python's library called NetworkX \cite{networkx}. NetworkX allow us to create a graph in witch we can storage suitable nodes
  that represent resplectively switches or hosts.
  This topological structure is updated live thanks to connection events and links discover offered by POX's libraries:
  \begin{itemize}
   \item \texttt{_handle_LinkEvent}: from openflow.discovery an event of this type gives us two types of information; a new link is added and 
   a still present link is expired;
   \item \texttt{_handle_ConnectionUp}: from core.openflow this event joins up when a new switch is added on the network and gives us
   some information of the switch, in particular its ports configuration;
   \item \texttt{_handle_ConnectionUp}: from core.openflow like the event explained above comes up when a switch is disconnected from the network;
   \item \texttt{_handle_port_status}: from core.openflow this event occures when on a particular switch there is a change about its ports, for example
   a new host has just been connected or the link is expired;\\
  \end{itemize}

@@

\subsubsection{Delay}
Monitoring round-trip time provides important insights for network troubleshooting and traffic engineering.
The  common  monitoring  technique  is  to  actively  send  probe packets  from  selected  vantage  points  hosts  or  middleboxes.
In traditional  networks, the control over the network routing is
limited,  making  it  impossible  to  monitor  every  selected  path. The  emerging  concept  of  Software  Defined  Networking  simplifies
network  control.  However,  OpenFlow,  the  common  SDN protocol,   does   not   support   RTT   monitoring   as   part   of   its
specification. In this experience we have tryed to emply an tecnic to escape this problem\cite{rtt}.
\newline The idea is to measure the time that occur from sending a packet from controller to a switch and the time in which the packet return from
an other switch.
\\
\newline First we have to discovery round trip time from controller to every single switch on our topology.
In order to do this on the controller we start a timer, we send a proper packet to the switch specifying the \texttt{OFPP_CONTROLLER - Send to the controller}
option and when the message return to the controller we stop the timer and calculate RTT for the switch.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt2.png}
 \caption{\emph{RTT between controller and OF switch}}
 \label{fig:topo}
\end{figure}
\newline
If we need to calculate the delay on a specific link, the rtt has to be calculated with the headed switch of this link.
\\
\newline
Now we create a proper message with unique id. number, let's start a timer \texttt{TM1} and then send this message to a first switch specifying the output port.
Then when the packet comes to the second switch it identify the particular message and sends to the controller that stops it's dedicated timer \texttt{TM2}.
\newline This part required a particular flow entry on the second switch in order to send to the controller only some particular packet.
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.70]{images/rtt1.png}
 \caption{\emph{Pck from controller to itself following a particular way.}}
 \label{fig:topo}
\end{figure}
\\
\newline At this point we can calculate the delay of the link.
\\
\\
 delay = (TM2 - TM1) - RTT1/2 - RTT2/2
\\
\\
Theoretically this method must works but in practice there are a lot of delay introduced from computational congestion on the virtualized controller.
What we observe is a estimated delay like a few seconds which is not the case, in fact if we made some \texttt{ping test} from one host to another host
in this SDN network we realize delay around 3ms.
\\
\\
The following picture shows the visual output that we can obtain from delay's graph. In this graph we storage only the core information;
\begin{figure}[!h]
 \centering
 \includegraphics[scale=0.26]{images/delay.png}
 \caption{\emph{Delay's graph.}}
 \label{fig:topo}
\end{figure}
